---
title: "Time Series Cross Validation"
author: "Alexander Häußer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series Cross Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "svg",
  fig.path = "figures/tscv-",
  fig.height = 5,
  fig.width = 7
)
```

The package `tscv` provides a collection of functions and tools for time series analysis and forecasting as well as time series cross-validation. This is mainly a set of wrapper and helper functions as well as some extensions for the packages `tsibble`, `fable` and `fabletools` that I find useful in the area of time series forecasting.

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ahaeusser/tscv")
```

```{r packages, message = FALSE, warning = FALSE}
# Load relevant packages
library(tscv)
library(tidyverse)
library(tsibble)
library(fable)
library(feasts)
Sys.setlocale("LC_TIME", "C")
```

## Data preparation

The dataset `elec_price` is a hourly `tsibble` with day-ahead electricity spot prices in [EUR/MWh] from the ENTSO-E Transparency Platform. The dataset contains time series data from 2019-01-01 to 2019-12-31 for 8 european bidding zones (BZN): 

* DE: Germany (including Luxembourg)
* DK: Denmark
* ES: Spain
* FI: Finland
* FR: France
* NL: Netherlands
* NO1: Norway 1 (Oslo)
* SE1: Sweden 1 (Lulea)

In this vignette, we will use only four time series to demonstrate the functionality of the package (the data set is filtered to the bidding zones Germany, France, Norway 1 and Sweden 1). You can use the function `clean_data()` to prepare the dataset for further usage. The function checks whether the input data are a valid tsibble or not (regular spaced in time and ordered). Furthermore, implicit missing values are turned into explicit missing values (existing missing values are left untouched). If the data are provided in wide format, they are gathered into long format. You can use the function `plot_line()` to visualize the four time series.

```{r clean_data}
# Prepare dataset
data <- elec_price %>%
  filter(BZN %in% c("DE", "FR", "NO1", "SE1")) %>%
  clean_data()

data

data %>%
  plot_line(
    x = Time,
    y = Value,
    color = BZN,
    facet = BZN,
    title = "Day-ahead Electricity Spot Price",
    subtitle = "2019-03-01 to 2019-03-15",
    xlab = "Time",
    ylab = "[EUR/MWh]",
    caption = "Data: ENTSO-E Transparency")
```


## Split data into training and testing

To prepare the dataset for time series cross-validation (TSCV), you can use the function `split_data()`. This function splits the data into training and testing (i.e. partitioning into time slices) for time series cross-validation. You can choose between `stretch` and `slide`. The first is an expanding window approach, while the latter is a fixed window approach. Furthermore, you can define the (initial) window size for training and testing via `n_init` and `n_ahead`, as well as the step size for increments via `n_skip`.

```{r split_data}
# Setup for time series cross validation
n_init <- 2400   # size for training window
n_ahead <- 24    # size for testing window (forecast horizon)
mode <- "slide"  # fixed window approach
n_skip <- 23     # skip 23 observations
n_lag <- 0       # no lag

data <- data %>%
  split_data(
    n_init = n_init,
    n_ahead = n_ahead,
    mode = mode,
    n_skip = n_skip,
    n_lag = n_lag)

data
```


The function `summarise_split()` provides a summary table of the partitioning into training and testing with the corresponding start and end (as date and index) for each split. This is very useful to identify specific splits by date. For example, if a holiday falls into a specific testing slice or not. 

```{r summarise_split}
# Summarize split into training and testing data
data %>% summarise_split()
```

## Training and forecasting

The data are splitted into training and testing slices and we are ready to forecast. Due to the sample size and computation time, only very simple benchmark methods are used. The functions `SMEDIAN` and `SMEAN` are extensions to the `fable` package. The function `SMEAN` is exactly the same as running a regression against seasonal dummy variables (`TSLM(value ~ season())`). I just added this function for convenience. Further forecasting methods are available (e.g. `TBATS()` and `DSHW()` from package `forecast` or `ELM()` and `MLP()` from package `nnfor`).

```{r train_models}
# Model training
data <- data %>%
  filter(split %in% c(1:100))

models <- data %>%
  filter(sample == "train") %>%
  model(
    sNaive = SNAIVE(Value ~ lag("week")),
    sMean = SMEAN(Value ~ lag("week")),
    sMedian = SMEDIAN(Value ~ lag("week")),
    "STL-Naive" = decomposition_model(STL(Value), NAIVE(season_adjust)))

models

# Forecasting
fcst <- models %>%
  forecast(h = n_ahead)

fcst

plot_forecast(
  fcst = fcst,
  data = data,
  include = 48,
  split = c(10, 11),
  title = "Day-ahead electricity spot price forecast",
  subtitle = "Rolling forecasts for splits 10 and 11",
  ylab = "[EUR/MWh]",
  caption = "Data: ENTSO-E Transparency, own calculation"
  )
```


## Evaluation of forecast accuracy

To evaluate the forecast accuracy, the function `error_metrics()` is used. You can define whether to evaluate the accuracy by `horizon` or by `split`. Several accuracy metrics are available:

* `ME`: mean error
* `MAE`: mean absolute error
* `MSE`: mean squared error
* `RMSE`: root mean squared error
* `MAPE`: mean absolute percentage error
* `sMAPE`: scaled mean absolute percentage error
* `MPE`: mean percentage error
* `MASE`: mean absolute scale error


### Forecast accuracy by forecast horizon
```{r accuracy_horizon}
# Estimate error metrics
metrics_horizon <- error_metrics(
  fcst = fcst,
  data = data,
  period = 168,
  by = "horizon")

metrics_horizon <- metrics_horizon %>%
  filter(metric %in% c("MAE", "MASE"))

metrics_horizon

# Visualize results
metrics_horizon %>%
  plot_error_metrics(
    title = "Evaluation of forecast accuracy by forecast horizon",
    subtitle = "Mean absolute scaled error (MASE)",
    xlab = "Forecast horizon (n-step-ahead)",
    caption = "Data: ENTSO-E Transparency, own calculation")
```


### Forecast accuracy by split
```{r accuracy_split}
# Estimate error metrics
metrics_split <- error_metrics(
  fcst = fcst,
  data = data,
  period = 168,
  by = "split")

metrics_split <- metrics_split %>%
  filter(metric %in% c("MAE", "MASE"))

metrics_split

# Visualize results
metrics_split %>%
  plot_error_metrics(
    title = "Evaluation of forecast accuracy by split",
    subtitle = "Mean absolute scaled error (MASE)",
    xlab = "Split",
    caption = "Data: ENTSO-E Transparency, own calculation")
```
